Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-05-03 overleaf/results.tex
Edits to results.

We used benchmarks from ABY and OPA~\cite{NDSS:DemSchZoh15,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{CCS:BDKKS18} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynth~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols). 
We did not use Gcd and Modular Exponentiation from~\cite{Ishaq:2019} as they do not benefit from vectorization according to our model. 
We ran as many routines from~\cite{CCS:BDKKS18} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
in our syntax and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
\cite{Anon_TR} describes what the benchmarks do.

To compare we run {\em both} versions with the maximal value of $N$ for which the non-vectorized version succeeds (typically {\tt N=128}) and we run {\em vec} with typically {\tt N=4096}.
Again, due to space constraints \cite{Anon_TR} gives the exact values.
% In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
%In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value for {\em both} and we fixed this value for these experiments. The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons).
% however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).

@@

A detailed summary of the effects of vectorization on various benchmarks is presented in \cref{table:metrics}. We show circuit evaluation times in \cref{fig:graph_all_eval_time}. In terms of observed benefits from vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include Convex Hull, Cryptonets Max Pooling, Minimal Points and Private Set Intersection. 
%These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol. 
2) {\it Medium:} these include Biometric Matching, DB Variance, Histogram, Inner Product, k-means Iteration and MNIST ReLU. These benchmarks have non-parallelizable phases e.g. the summing phase of Inner Product. Still, most computation is parallelizable and it results in speedup from 5x to 25x in BMR, and 2x to 25x in GMW protocol. 3) {\it Low:} these include the Database Join and the regular expression benchmarks (Count 102, Count 10, Longest 102 and Max distance between symbols). There is less parallelizable computation in these programs, thus the speedup is lower. We see a speedup from 1.1x to 2x in BMR. In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x. However, Longest 102 and Max distance between symbols suffer a slowdown of 0.5x. There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, a there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized. We observed that transformation to vectorized code increased multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW. The cause of the increase is not clear --- we conjecture that MOTION performs optimizations over the non-vectorized loop body that decreases depth; also, EQ is relatively inexpensive in Boolean GMW and BMR compared to ADD and MUL, which also de-emphasizes the benefit of vectorization. We propose a simple heuristic although we do leave all the benchmarks in the table: if the transformation increases circuit depth beyond some threshold (e.g. more than 10\% of the original circuit), we can reject the transformation. %Nevertheless, we show these graphs here for the sake of completeness and to highlight that vectorization is not always reduce run time. 
Note that in some settings it may still be desirable to vectorize e.g. in data constrained environments as communication as well as all other metrics, is reduced for all benchmarks. (The report presents plots of for all metrics~\cite{Anon_TR}.)%As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).

@@

MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to
limitations of our implementation. Recall that Phase 3 of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that static analysis 
will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above slightly more complex code. 
However, because AST analysis is difficult it does not appear to detect the accumulation.
%(unfortunately, we ran out of time and did not experiment further).
%Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.

@@

HyCC~\cite{CCS:BDKKS18} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ as its focus is mixing. %; it does provide a lot of data on mixed protocol executions.
We estimate we are about an order of magnitude slower, which is likely due to the same issue --- the computation of $\mathit{min}$ can be optimized.

>>>>

We used benchmarks from ABY and OPA~\cite{NDSS:DemSchZoh15,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{CCS:BDKKS18} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynt~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols). 
We did not use Gcd~\cite{Ishaq:2019} and Modular Exponentiation from~\cite{NDSS:DemSchZoh15} as they do not benefit from vectorization according to our model. 
We ran as many routines from~\cite{CCS:BDKKS18} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
in our syntax and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
\cite{Anon_TR} describes what the benchmarks do.

To compare we run {\em both} versions with the maximal value of $N$ for which the non-vectorized version succeeds (typically {\tt N=128}). We also run {\em vec} with larger values, 
typically {\tt N=4096}. Again, due to space constraints \cite{Anon_TR} gives the exact values.
% In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
%In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value for {\em both} and we fixed this value for these experiments. The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons).
% however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).

@@

A detailed summary of the effects of vectorization on various benchmarks is presented in \cref{table:metrics}. We show circuit evaluation times in \cref{fig:graph_all_eval_time}. In terms of observed benefits from vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include Convex Hull, Cryptonets Max Pooling, Minimal Points and Private Set Intersection. 
%These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol. 
2) {\it Medium:} these include Biometric Matching, DB Variance, Histogram, Inner Product, k-means Iteration and MNIST ReLU. These benchmarks have non-parallelizable phases e.g. the summing phase of Inner Product. Still, most computation is parallelizable and it results in speedup from 5x to 25x in BMR, and 2x to 25x in GMW protocol. 3) {\it Low:} these include the Database Join and the regular expression benchmarks (Count 102, Count 10, Longest 102 and Max distance between symbols). There is less parallelizable computation in these programs, thus the speedup is lower. We see a speedup from 1.1x to 2x in BMR. In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x. However, Longest 102 and Max distance between symbols suffer a slowdown of 0.5x. There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized. We observed that transformation to vectorized code increased multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW. The cause of the increase is not clear --- we conjecture that MOTION performs optimization over the non-vectorized loop body that decreases depth; also, EQ is relatively inexpensive in Boolean GMW and BMR compared to ADD and MUL, which also de-emphasizes the benefit of vectorization. We propose a simple heuristic: if the transformation increases circuit depth beyond some threshold (e.g. more than 10\% of the original circuit), we can reject the transformation. Nevertheless, we show these benchmarks here for the sake of completeness and to highlight that vectorization does not always reduce run time. Note that in some settings it may still be desirable to vectorize e.g. in data constrained environments as communication as well as all other metrics, is reduced for all benchmarks. (The extended version presents plots of for all metrics~\cite{Anon_TR}.)%As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).

@@

MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to
limitations of our implementation. Recall that Phase 3 of Vectorization in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that static analysis 
will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above simpler code. 
However, because AST analysis is difficult, it does not appear to detect the accumulation.
%(unfortunately, we ran out of time and did not experiment further).
%Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.

@@

HyCC~\cite{CCS:BDKKS18} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ (as its focus is mixing). %; it does provide a lot of data on mixed protocol executions.
We estimate we are about an order of magnitude slower, which is likely due to the same issue --- the computation of $\mathit{min}$ can be optimized.
